---
title: "Individual-level panel data confirm the widening gender gap in publishing rates during COVID-19"
knit: (function(input_file, encoding) {
  out_dir <- "docs";
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, "index.html"))})
author: "Emil Bargmann Madsen, Mathias Wullum Nielsen, Josefine Bjørnholm, Reshma Jagsi, and Jens Peter Andersen"
date: "04/01/2022"
output: 
  html_document:
    theme: readable
    toc: true
    toc_float: 
      collapsed: false
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE,echo=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

list_of_packages <- c("tidyverse", "extrafont", "scales", "broom", "patchwork", "kableExtra", "fixest", "jtools", "margins", "sandwich", "MASS", "patchwork", "rio")
lapply(list_of_packages, require, character.only = TRUE)

```

# Introduction

This is a document containing the tables and figures for a study on the gendered impact of the COVID-19 pandemic on publication productivity of individual researchers by Emil Bargmann Madsen, Mathias Wullum Nielsen, Josefine Bjørnholm, Reshma Jagsi, and Jens Peter Andersen. The source code can be found at: https://github.com/emilbargmann/covid_update/


# Main document figures

## Figure 1

```{r,fig.cap="Average publication output by gender and year. Differences are in percentages of average publication rates in 2019. Results are presented for full and fractionalized publication counts.",fig.width = 8, fig.height = 4}

source("figures/figure_1_gen.R")
figure_1

```

## Figure 2

```{r,fig.cap="Dynamic effects of the COVID-19 pandemic on women’s and men’s publication productivity. Panel A shows the estimated average gender difference in publication rates by year. Panel B shows the predicted publishing rates for men and women authors. Dashed, colored, lines represent the counterfactual trend for women if they had similar 2019-2020 trajectories as men. The difference between the dashed line and the straight line in Panel B specifies the average treatment effect for women. Point estimates are reported with 99.9% confidence bounds, with robust standard errors clustered at the individual-author level. Regression tables underlying these results can be found in Supplementary Tables S1-2. Corresponding regressions using Poisson and negative binomial models are in Supplementary Tables S3-4.", fig.width = 8,fig.height = 4}

source("figures/figure_2_gen.R")
figure_2

```

## Figure 3

```{r,fig.cap="Dynamic effects of the COVID-19 pandemic on the average gender gap in annual publishing rates, by career age. Upper panels show the estimated average gender difference in full-count publication rates by year for early career and mid-career researchers. Lower panels show men’s and women’s predicted full-count publication rates per year by author status (early-career vs. mid-career researcher). Point estimates are reported with 99.9% confidence bounds and robust standard errors clustered at the individual-author level. Regression tables underlying these results can be found in Supplementary Tables 9-12. The corresponding analysis with fractional counts can be found in Supplementary Figure 4.", fig.width = 8,fig.height = 8}

source("figures/figure_3_gen.R")
figure_3

```

## Figure 4

```{r,fig.cap="Stratified effects of the COVID-19 pandemic on the average gender gap in annual publishing rates. Panel A shows the estimated average gender difference in publication rates by year. Panel B shows the predicted publishing rates for men and women authors. In each panel, scientists are divided into strata according to their total number of publications in the period 2016-2020. The difference between the thinner, dashed line with the black circle in 2020 and the thicker, dashed line in Panel B specifies the average treatment effect for women. Point estimates are reported with 99.9% confidence bounds and robust standard errors clustered at the individual-author level.", fig.width = 8,fig.height = 4}

source("figures/figure_4_gen.R")
figure_4

```

## Figure 5

```{r,fig.cap="Gender differences in full and fractionalized publication productivity by country, 2019 vs. 2020. 2020 gender differences in full publications counts relative to 2019, across 30 countries comprising 90% of authors in our sample. Difference-in-differences estimate from Supplementary Table S13. 99% confidence intervals based on robust clustered standard errors are shown. Histograms show the distribution of authors who mainly publish within a given discipline, and orange and green lines and bands show the overall treatment effect on the sample and the counterfactual sample from regression tables S1 and S7 in the supplementary material (ATT is the Average Treatment effect on the Treated).", fig.width = 8,fig.height = 8}

source("figures/figure_5_gen.R")
figure_5

```

## Figure 6

```{r,fig.cap="Gender differences in full and fractionalized publication productivity by discipline, 2019 vs. 2020. 2020 gender differences in full publications counts relative to 2019, across the four disciplines comprising in our sample. Difference-in-differences estimate from table S14 in the supplementary material. 99% confidence intervals based on clustered standard errors are shown. Histograms show the distribution of authors who mainly publish within a given discipline, and orange and green lines and bands show the overall treatment effect on the sample and the counterfactual sample from regression tables S1 and S7 in the supplementary material. Panel B shows the distribution of authors per discipline in deciles of total publications over the time period.", fig.width = 7,fig.height = 7}

source("figures/figure_6_gen.R")
figure_6

```

## Figure 7

```{r,fig.cap="", fig.width = 8,fig.height = 7}

source("figures/figure_7_gen.R")
figure_7

```

# Supplementary figures

## Figure S1

```{r,fig.cap="Dynamic effects of the COVID-19 pandemic on women’s and men’s fractional-count output. Panel A shows the estimated average gender difference in fractional-count publication rates by year. Panel B shows the predicted fractional-count publishing rates for male and female authors. Dashed, colored, lines represent the counterfactual trend for women if they had similar 2019-2020 trajectories as men. The difference between the dashed line and the straight line in Panel B specify the average treatment effect for women. Point estimates are reported with 99.9% confidence bounds, with robust standard errors clustered at the individual-author level. Regression tables underlying these results can be found in Supplementary table S2.", fig.width = 8,fig.height = 4}

source("figures/figure_S1_gen.R")
figure_S1

```

## Figure S2

```{r,fig.cap="Dynamic effects of the COVID-19 pandemic on the average gender gap in annual publishing rates, by career age. The upper panels show the estimated average gender difference in fractional-count publication rates by year for early career and senior researchers. The lower panels show men’s and women’s predicted fractional-count publication rates per year by author status (early-career vs. senior researcher). Point estimates are reported with 99.9% confidence bounds, with robust standard errors clustered at the individual-author level. Regression tables underlying these results can be found in SI Appendix Tables S11-12.", fig.width = 8,fig.height = 8}

source("figures/figure_S2_gen.R")
figure_S2

```

## Figure S3

```{r,fig.cap="Gender differences in publication productivity for at counterfactual sample. Black lines show average marginal effects of female author gender each year, and colored lines show predicted publication counts (both full and fractionalized) for male and female authors. Predicted differences and counts are based on the difference-in-differences estimates from table S7-8 in the supplementary material. 99.9 % confidence intervals based on clustered standard errors are shown.", fig.width = 8,fig.height = 8}

source("figures/figure_S3_gen.R")
figure_S3

```

## Figure S4

### Subfigure 1
```{r,fig.cap="Gender differences in full- and fractionalized-count publication productivity, 2019 vs. 2020, across four different lockdown severity indicators. All indicators are summed values as stipulated in Equation 3. 99% confidence intervals based on clustered standard errors are shown.", fig.width = 8,fig.height = 8}

source("figures/figure_S4_gen.R")
figure_S4_1

```

### Subfigure 2
```{r,fig.cap="Gender differences in full- and fractionalized-count publication productivity, 2019 vs. 2020, across three different lockdown severity indicators. All indicators are counts of the maximum values as stipulated in Equation 4. 99% confidence intervals based on clustered standard errors are shown.", fig.width = 8,fig.height = 8}

source("figures/figure_S4_gen.R")
figure_S4_2

```

# Supplementary tables


## Table S1
```{r}

source("tables/table_S1.R")
table_S1

```

## Table S2
```{r}

source("tables/table_S2.R")
table_S2

```

## Table S3
```{r}

source("tables/table_S3.R")
table_S3

```

## Table S4
```{r}

source("tables/table_S4.R")
table_S4

```

## Table S5
```{r}

source("tables/table_S5.R")
table_S5

```

## Table S6
```{r}

source("tables/table_S6.R")
table_S6

```

## Table S7
```{r}

source("tables/table_S7.R")
table_S7

```

## Table S8
```{r}

source("tables/table_S8.R")
table_S8

```

## Table S9
```{r}

source("tables/table_S9.R")
table_S9

```

## Table S10
```{r}

source("tables/table_S10.R")
table_S10

```

## Table S11
```{r}

source("tables/table_S11.R")
table_S11

```

## Table S12
```{r}

source("tables/table_S12.R")
table_S12

```

## Table S13
```{r}

source("tables/table_S13.R")
table_S13

```

## Table S14
```{r}

source("tables/table_S14.R")
table_S14

```